Robust Control of a Fluxonium Qubit

### 2020-11-03 paper/main.tex
updates to paper

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity operations on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
  but they require exquisite calibration to be performant.
  In this work we employ robust trajectory optimization techniques
  to achieve high fidelity gates for parameter spreads on the order of $1\%$.
  We propose a method that takes advantage of deviant parameter derivative information while maintaining
  computational efficiency by utilizing mixed-mode differentation.
  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
  the Lindblad master equation, which increases the computational complexity
  of the optimization algorithm. We propose a computationally efficient metric
  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
  We apply these techniques to a fluxonium qubit with realisitc experimental constraints
  and noise values. We are able to reduce gate errors by orders of magnitudes in the presence
  of deviant parameters, achieving super-linear scaling for gate robustness as a function of the gate duration.
  We use the same techniques to mitigate coherent errors due to time-varying magnetic flux noise
  by two orders of magnitude. Additionally, we mitigate the gate error arising from longitudinal
  relaxation by a factor of $5$. Further, we find that the numerical techniques we employ
  are able to perform phase gates in arbitrary times, which is critical for operating multi-qubit
  experiments in the lab frame.
\end{abstract}

@@

%% S1
\section{Introduction}
%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
%% P1 - general field, lots of citation. cite reviews, original qoc.
%% P2 - exhaustively find all papers, what people have done to address the issues
%% P3 - what we do is. maybe add a rant about how
%% QOC is classical control theory but we've ignored that field
%% P4 - outline
The field of quantum optimal control (QOC) is concerned
with efficiently and accurately manipulating quantum systems.
Early QOC techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}. QOC techniques are used to
control a system such that a set of optimization objectives are maximally fulfilled.
For quantum computation, relevant objectives include achieving high fidelity
gates and adhering to experimental constraints.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.
It is desirable to make the optimization aware of experimental errors
such as noise, system model errors, and control errors, to
improve experimental performance on the optimization objectives.
The field of robust control, deriving from classical control theory, has developed
techniques to account for these errors in optimization (citations appreciated).
In this work we employ robust control techniques to mitigate
releastic decoherence and systematic errors that arise when controlling
a superconducting fluxonium qubit.

Numerical techniques for QOC have been remarkably successful in
state preparation, designing high fidelity gates, and determining
optimized pulse sequences (citations needed). Many popular numerical techniques
are iterative, indirect shooting
methods \cite{leung2017speedup, goer2019krotov, abdelhafez2019gradient,
  machnes2015gradient}.
An initial guess for the control parameters is used to forward-simulate
the quantum state trajectory under the time-dependent Schroedinger equation. Then, gradients
with respect to the controls are computed for an objective evaluated on
the evolved quantum state. The gradients are used to update
the control parameters and the procedure is repeated iteratively.
A convenient way to enforce constraints in this framework is
to employ control filtering or project gradients onto the constraint manifold.
In this work we perform optimization using ALTRO, which combines
an iterative indirect shooting method with the augmented Lagrangian
method \cite{howell2019altro}. The augmented Lagrangian method
allows us to enforce arbitrary, simultaneous constraints and achieve
fast convergence without being restricted to the constraint
manifold.

QOC techniques have been effective in mitigating the effects
of systematic errors. Typically, the model used in optimization
differs from the experimental system due to parameter drift, the inability
to realize control parameters exactly, and noise.
These systematic errors cause the trajectory of the quantum system
in the experiment to deviate from that predicted by the optimization.
A set of controls is said to be robust if the experimental state trajectory
it produces is insensitive to systematic errors.
Considering the dynamical and geometric phases of quantum state
evolution has led to analytic methods for achieving
robustness to dynamical errors and mitigating pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
We mitigate systematic errors numerically by employing
open loop optimization. We draw on robust trajectory optimization
techniques to encode systematic errors in optimization objectives.
We consider three
techniques. The first is the sampling method, which has been applied
previously in the context of QOC
\cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
The second is the unscented sampling
method, which derives from the unscented transformation used
for non-linear Kalman filtering
\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
We propose a third method, which we call the derivative method.
This method takes advantage of derivative information of systematic
errors with respect to the state evolution. We make this method efficient
by employing mixed-mode differentiation.

Further, many analytic and numerical techniques have been developed to mitigate decoherence.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
decoherence due to longitudinal relaxation and pure dephasing
\cite{huang2020engineering, mundada2020floquet}.
Numerical techniques have mitigated decoherence by evolving the
system using master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
In this work we take advantage of the known dependence of the
quantum system's controls on longitudinal relaxation. We encode
the decoherence in an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.

We compare our techniques using the heavy fluxonium qubit presented in
in \cite{zhang2020universal}. This quantum system is an exciting
platform for quantum computation due to its long longitudinal relaxation
and pure dephasing times at the flux frustation point. This system is also
efficient to simulate due to its accurate two-level approximation.
We emphasize that the numerical techniques we employ extend to arbitrary quantum systems,
not only the one we study here. We describe experimentally realistic constraints for this system
and map them to the ALTRO framework. Next, we
outline a method for making the optimization aware of longitudinal
relaxation. We achieve a factor of 5 increase in longitudinal relaxation times
over the baseline gate set. Then we present three methods for achieving
robustness to systematic errors. We find that we are able to
decrease the gate error arising from systematic errors super-linearly
in the gate duration, achieving orders of magntidue reduction in gate errors.
We also find that the numerical techniques we employ are able to produce
fixed phase gates in arbitrary times, a method that could be used
to control multi-qubit systems in the lab frame. Finally,
we employ the robust control techniques to mitigate time-varying
magnetic flux noise. We achieve a two order of magnitude improvement
in gate errors for extended computations, an improvement critical
for noisy, intermediate-scale quantum (NISQ) applications.

@@

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines.
%% TODO: See ref. 28 in Helin paper
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \ref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.

The optimization is performed over the second derivative of the flux amplitude
$\frac{d^{2}}{dt^{2}} a$ which is contained in the
augmented control vector. The first derivative
$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$.
The equality and inequality constraints on $a$ are handled
with a bound constraint.
%% TODO: probably want to list equations for the constraints

In addition to obeying experimental constraints
and achieving low simulated gate errors, we desire
to make our gates perform well in the presence of decoherence.
Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
(see Appendix).
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.

Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
There is a tradeoff between the two decoherence processes. In the case of white
noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
is constant \cite{huang2020engineering}.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
via the flux amplitude and employ robust control techniques to mitigate
pure dephasing.

@@

%% S5
\section{Robustness to Static Parameter Deviations}
We have formulated the quantum optimal control
problem as an open loop optimization problem, i.e.
feedback from the experiment is not incorporated in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. To mitigate
these errors we employ robust control techniques
to make the state evolution insensitive to Hamiltonian
parameter deviations. As an example
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
We consider three robust control techniques.
The first is the sampling method, which has
been preposed previously in the context of QOC
\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
sampling method, multiple states are evolved under distinct deviant dynamics
to capture the effect of parameter deviations. We also
study the unscented sampling method which uses the unscented
transformation to accurately propagate a distribution
representing the uncertainty in an evolving state
due to a parameter deviation.
The unscented transformation was designed for nonlinear Kalman
filtering and is frequently utilized in robust control
\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
Finally, we propose the derivative method. Derivative
information encoding the sensitivity of the
state trajectory with respect to the deviant parameter
is used to penalize state trajectory deviations.

@@

%% S7
\section{Conclusion}
We have proposed some schemes and they work well.

>>>>

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity gates on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have proven effective in experimentally
  realizing high fidelity gates, but they require exquisite calibration to be performant.
  We apply robust trajectory optimization techniques to suppress gate errors arising from system
  model deviations and noise.
  We propose a method that takes advantage of deviant parameter derivative information while maintaining
  computational efficiency by utilizing mixed-mode differentation.
  Additionally, completely modeling decoherence effects due to longitudinal relaxation requires
  integrating the Lindblad master equation, which is computationally expensive.
  We propose a computationally efficient metric
  and utilize time-optimal control to achieve high fidelity gates in the presence of longitudinal relaxation.
  We demonstrate these techniques numerically on a fluxonium qubit with realisitc experimental parameters and constraints,
  achieving orders of magnitude gate error reductions from our baseline gate set.
  %% TODO: maybe add arbitrary phase bit
\end{abstract}

@@

%% S1
\section{Introduction}
%% textwidth: \printinunitsof{in}\prntlen{\textwidth}
%% linewidth: \printinunitsof{in}\prntlen{\linewidth}
%% P1 - general field, lots of citation. cite reviews, original qoc.
%% P2 - exhaustively find all papers, what people have done to address the issues
%% P3 - what we do is. maybe add a rant about how
%% QOC is classical control theory but we've ignored that field
%% P4 - outline
Quantum optimal control (QOC) techniques are a class of optimization
algorithms for accurately and efficiently manipulating quantum systems.
Early techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
circuits \cite{heeres2017implementing,
  huang2020engineering, leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}. QOC techniques aim to control the system
such that a set of objectives are optimally satisfied.
For quantum computation relevant objectives include achieving high fidelity
gates and adhering to experimental constraints.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.
Experimental errors may cause the system evolution to deviate from that predicted in
optimization, hampering performance.
The field of classical control theory has developed robust control techniques
to encode experimental errors in optimization objectives, improving
performance on the experiment (citations appreciated).
In this work we employ robust control techniques to mitigate
releastic decoherence and systematic errors that arise when controlling
a superconducting fluxonium qubit.

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates (citations needed). In addition to the high
simulated fidelities which can be obtained with these techniques, it is desirable for the
resultant quantum state trajectory to be insensitive to experimental errors
such as parameter drift, noise, and finite control resolution. Controls that produce
quantum state trajectories that are insensitive to these effects are said to be robust.
In line with this objective, recent work has sought to mitigate systematic errors
and decoherence present in the experiment.
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for achieving
robustness to systematic errors and mitigating pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
decoherence due to longitudinal relaxation and pure dephasing
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

In this work we employ state of the art trajectory optimization techniques
to mitigate systematic errors and decoherence. We frame the QOC problem as an
open loop optimization problem, optimizing the controls to make the
quantum state trajectory robust to systematic errors and decoherence.
We consider three robust trajectory optimization techniques.
The first is the sampling method, which has been applied
previously in the context of QOC
\cite{carvalho2020error, reinhold2019controlling, rembold2020introduction}.
The second is the unscented sampling
method, which derives from the unscented transformation used
for non-linear Kalman filtering
\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
We propose a third method, which uses derivative information
of parameter deviations with respect to the quantum state evolution.
We make this method efficient by employing mixed-mode differentiation.
Additionally, we take advantage of the known dependence of the
system's controls on longitudinal relaxation. We encode
the decoherence in an efficient optimization objective that does
not pay the increased computational cost of integrating a master equation.
We perform these optimization using the ALTRO framework \cite{howell2019altro}.
Many popular numerical techniques
for QOC use gradient-based control update procedures, which are readily amenable to control filtering
or constraint manifold projection techniques to enforce constraints \cite{leung2017speedup,
  goer2019krotov, abdelhafez2019gradient, machnes2015gradient}.
The ALTRO method we employ combines
an iterative indirect shooting method with the augmented Lagrangian
method. The augmented Lagrangian method
allows us to enforce arbitrary, simultaneous constraints and achieve
fast convergence without being restricted to the constraint
manifold.

We compare our techniques using the heavy fluxonium qubit and assocaited analytic gates
presented in \cite{zhang2020universal}. This quantum system is an exciting
platform for quantum computation due to its long longitudinal relaxation
and pure dephasing times at the flux frustation point. This system is also
efficient to simulate due to its accurate two-level approximation.
We emphasize that the numerical techniques we employ extend to arbitrary quantum systems,
not only the one we study here. We describe experimentally realistic constraints for this system
and map them to the ALTRO framework. Next, we
outline a method for making the optimization aware of longitudinal
relaxation. We achieve a factor of 5 increase in longitudinal relaxation times
over the baseline gate set. Then we present three methods for achieving
robustness to static parameter deviations. We find that we are able to
decrease the gate error arising from parameter deviations super-linearly
in the gate duration, achieving orders of magntidue reductions in single gate errors.
We also find that the numerical techniques we employ are able to produce
fixed phase gates in arbitrary times, a method that could be used
to control multi-qubit systems in the lab frame. Finally,
we employ the robust control techniques to mitigate time-varying
magnetic flux noise. We achieve a two order of magnitude improvement
in gate errors for extended computations, an improvement critical
for noisy, intermediate-scale quantum (NISQ) applications.

@@

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines.
%% TODO: See ref. 28 in Helin paper
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \ref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\frac{d^{2}}{dt^{2}} a$ which is contained in the
augmented control vector. The first derivative
$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$.
The equality and inequality constraints on $a$ are handled
with a bound constraint.
%% TODO: probably want to list equations for the constraints

In addition to obeying experimental constraints
and achieving low simulated gate errors, we desire
to make our gates perform well in the presence of decoherence.
Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
(see Appendix).
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.
Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
There is a tradeoff between the two decoherence processes. In the case of white
noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
is constant \cite{huang2020engineering}.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
by tuning the flux amplitude and employ robust control techniques to mitigate
pure dephasing.

@@

%% S5
\section{Robustness to Static Parameter Deviations}
We have formulated the quantum optimal control
problem as an open loop optimization problem, i.e.
feedback from the experiment is not incorporated in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. To mitigate
these errors we employ robust control techniques
to make the state evolution insensitive to Hamiltonian
parameter deviations. As an example
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
We consider three robust control techniques.
The first is the sampling method, which has
been proposed previously in the context of QOC
\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
sampling method, multiple states are evolved under distinct deviant dynamics
to capture the effect of parameter deviations. We also
study the unscented sampling method which uses the unscented
transformation to accurately propagate a distribution
representing the uncertainty in an evolving state
due to a parameter deviation.
The unscented transformation was designed for nonlinear Kalman
filtering and is frequently utilized in robust control
\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
Finally, we propose the derivative method. Derivative
information encoding the sensitivity of the
state trajectory with respect to the deviant parameter
is used to penalize state trajectory deviations.

@@

%% S7
\section{Conclusion}
In conclusion, we have demonstrated techniques for achieving robustness to systematic
errors and mitigating decoherence on a quantum system using state-of-the-art trajectory
optimization. We have proposed a scheme for mitigating longitudinal relaxation with time-optimal
control and an efficient optimization metric that comes at a constant computational cost as
opposed to integrating a master equation which scales quadratically with
the dimension of the Hilbert space.
We have proposed the derivative method for robust control which achieves
super-linear gate error reductions in the gate duration for the problem we studied here.
We have shown that robust control techniques can be used to mitigate decoherence due
to 1/$f$ flux noise, a dominant source of coherent errors for flux controlled qubits.
The numerical techniques we have studied can be used to perform phase gates in arbitrary times,
which will be critical for synchronizing multi-qubit systems. Our
techniques significantly improve the state-of-the-art for controlling quantum systems subject
systematic errors and decoherence, and will be used in efforts to lower gate errors
for fault-tolerant quantum computing and NISQ applications.
